---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# trans

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/trans)](https://CRAN.R-project.org/package=trans)
<!-- badges: end -->

> [!WARNING]  
> This package is in early development and not yet ready for use

The trans R package provides functions for basic 3d & multidimensional transformation of points, vectors, and planes.

## Installation

You can install the development version of trans like so:

``` r
if (!require("remotes"))
    install.packages("remotes")

remotes::install_github("selkamand/trans")
```

## Quick Start

### Rotate a vector around an axis (Rodrigues)

```{r}
library(trans)
library(rgl)

v <- c(1, 0, 0)
axis <- c(0, 0, 1)
angle <- pi/2  # 90 degrees in radians

rotate_vector_around_axis(v, axis, angle)
```

### Rotate a vector to align with a target direction

```{r}
v <- c(2, -3, 4)    # any vector
target <- c(1,  1, 1)    # direction we want

v_rot <- rotate_vector_to_align_with_target(v, target)
sqrt(sum(v^2))     # original length
sqrt(sum(v_rot^2)) # preserved length
```

### Rotate a vector into a plane (not projection)

```{r}
v <- c(1, 2, 3)
n <- c(0, 1, -1)     # plane normal

v_in_plane <- rotate_vector_into_a_plane(v, n)
# v_in_plane lies in the plane; ||v|| is preserved (unlike a projection)
```

### Project Vectors

```{r}
# Vector projection of a onto b
project_vector_into_vector(c(2, 3, 4), c(1, 0, 0))

# Scalar projection (signed length of a along b)
compute_scalar_projection(c(2, 3, 4), c(1, 0, 0))
```

### Translate positions

```{r}
translate_position_by_vector(c(1, 2, 3), c(0.5, 0, -1))
#> c(1.5, 2, 2)

translate_position_in_direction(c(0, 0, 0), c(1, 1, 0), 10)
#> approximately c(7.071068, 7.071068, 0)

compute_translation_vector(c(1, 2, 3), c(4, 6, 3))
#> c(3, 4, 0)
```


### Measuring

```{r}
a <- c(1, 0, 0)
b <- c(0,1,0)

measure_angle_between_vectors(a, b)

a %angle% b
```

### Conversion

```{r}
radians_to_degrees(pi)
degrees_to_radians(180)
```


### Plane representation conversions

Planes can be represented in two equivalent forms:

1. **Point–normal form**  
   Defined by a 3D point `point` that lies on the plane and a normal vector `normal`.  
   Equation:  
   \[
   \hat{n} \cdot (x - p) = 0
   \]

2. **Normal–offset (Hesse) form**  
   Defined by a **unit** normal vector `normal` and a **signed offset** (distance) `offset`
   from the origin along that normal.  
   Equation:  
   \[
   \hat{n} \cdot x = s
   \]
   where `s` = `offset`.

These forms can be converted with the following functions:

```{r}
# Point–normal -> Normal–offset
x <- convert_plane_point_normal_to_normal_offset(normal = c(0,0,2), point = c(2,3,5))
x$normal  # c(0,0,1)
x$offset  # 5

# Normal–offset -> Point–normal (closest point to origin)
convert_plane_normal_offset_to_point(normal = c(0,0,1), offset = 5)
```


